### 2016-11-28 cover_letter.tex

\documentclass[a4paper,12pt]{article} 
\usepackage{times}
\usepackage[top=1in,bottom=1in,left=1in,right=1in,nohead]{geometry}
\usepackage{graphics}
\usepackage{hyperref}

\begin{document}
\renewcommand{\rmdefault}{ptm}
\pagestyle{empty} 
\small

\setlength{\parindent}{0mm} 
\setlength{\parskip}{5mm}

\begin{flushright}
\today
\end{flushright}

\vfill
\emph{To:}\newline
Bettina Gr\"un, Edzer Pebesma, Achim Zeileis\newline
Editors-in-Chief\newline
Journal of Statistical Software

\emph{From:}\newline
Marcus W. Beck\newline
US Environmental Protection Agency\newline
beck.marcus@epa.gov, 850-934-2480\vspace{0.1in}

Enclosed please find my revised manuscript, entitled `NeuralNetTools: Visualization and Analysis Tools for Neural Networks'.  This revision follows a minor acceptence as a research article in JSS.

I appreciate the insightful comments provided by the reviewer and I have made every effort to address the concerns.  In particular, I have improved the applied example by demonstrating the interpretation of a simple neural network model using functions from the package. This includes addition to the text and Figure 7.  I have also fixed any remaining issues with reproducibility in the manuscript code and supplementary material.  Please see all of my responses to the reviewer comments included with my revision.

This revision includes a manuscript PDF created with the JSS \LaTeX\ template (nnt\_manu.pdf), my response to the reviewer comments (review\_response.pdf), an R script of the manuscript code (nnt\_manu.R), a .tex file for the manuscript text (nnt\_manu.tex), a .bbl file with bibliographic information (refs.bbl), all pdf files used for figures (in fig folder), and supporting JSS \LaTeX\ template files.  Links to the package source code are at the bottom of this letter. The package is licensed uncer CC0.

I greatly appreciate the opportunity to publish my work in JSS.  Please feel free to contact me directly should additional information be needed.

\vspace{0.1in}

\hspace{4.5in}Respectfully,

\hspace{4.5in}Dr. Marcus W. Beck

{\it Package sources}

Github (dev): \url{https://github.com/fawda123/NeuralNetTools}

CRAN (stable): \url{https://cran.r-project.org/web/packages/NeuralNetTools/index.html}

\vfill
\end{document}